---
title: "Chapter 5 Homework"
output: html_notebook
---

### Malaria medication adherence

In `malaria.csv` we have the results from an experiment that provided text alerts to encourage adherence to antimalarial medication (taking the meds) in Ghana.  The data are taken from the paper at https://doi.org/10.7910/DVN/M4LY6C (*Raifman, Julia RG, et al. "The impact of text message reminders on adherence to antimalarial treatment in northern Ghana: a randomized trial." PloS one 9.10 (2014)*). 

The `treated` variable is whether or not they get the text alerts, the `adhere` variable is whether they (mostly) adhered to the medication protocol, and the `age` and `male` variables tell us the patient's age and whether they are male. 

<br>
<br>

**Calculate a 95% interval for the average treatment effect of text reminders on adherence probability.**

```{r}
d <- read.csv("malaria.csv")
fit1 <- glm( adhere ~ treated, data=d)
( stats1 <- summary(fit1)$coef["treated",] )
# confidence interval
stats1[1] + c(-1,1)*1.96*stats1[2]
```

The interval barely overlaps with zero, so this is close but not quite significant at the 0.05 level.

Alternatively, you can bootstrap.
```{r}
library(parallel)
library(boot)
getATE <- function(data, ind){
  coef(glm(adhere ~ treated, data=data[ind,]))["treated"]
}
set.seed(1)
(boot1 <- boot(d, getATE, 1000, 
               parallel="snow", ncpus=detectCores()))
## percentile bootstrap
quantile(boot1$t, c(.025,.975))
## bias adjusted
quantile(2*boot1$t0 -boot1$t, c(0.025,.975))
```
Similar results here to the theoretical CI derived from the glm output.

**Do you have any evidence that the treatment selection is driven by the patient's age or sex?**

This is asking if the age and `male` covariates are predictive of treatment status.  Running the regression below indicates that they are not.  

```{R}
summary(glm(treated ~ age*male, data=d))
```

<br>

**Break the age into ten year chunks (1-10, 11-20, etc) and estimate a heterogeneous treatment effects model where the treatment effect depends upon age category and sex.  What is the 95% CI for the CATE for a 25 year old female?**

First, use `cut` to break the age into chunks.
```{r}
max(d$age)
d$agegrp <- cut(d$age,10*(0:9))
levels(d$agegrp)
```
Then, run a regression model.  You could use linear or logistic regression; we will use logistic regression since the response is binary.
```{r}
fit2 <- glm(adhere ~ treated*agegrp*male, data=d)

getCATE <- function(data, ind){
  fit <- glm(adhere ~ treated*agegrp*male, data=data[ind,])
  predict(fit, data.frame(treated=1, male=0, agegrp="(20,30]")) - 
    predict(fit, data.frame(treated=0, male=0, agegrp="(20,30]"))
  }
set.seed(1)
(boot2 <- boot(d, getCATE, 1000, 
               parallel="snow", ncpus=detectCores()))
## percentile bootstrap
quantile(boot2$t, c(.025,.975))
## bias adjusted
quantile(2*boot2$t0 -boot2$t, c(0.025,.975))
```

Note that the warning is because we include factor levels that are not identified.  You can ignore it.

<br>
**Simulate data from a regression discontinuity design and recover an estimate of your treatment effect.**

Our simulator has two linear models.  The treatment effect at r=0 is 3-1=2 (difference in intercepts)
```{r}
fcntrl <- function(r){ 1 +10*r + rnorm(length(r), sd=5) }
ftreat <- function(r){ 3 +5*r + rnorm(length(r), sd=5)}
```

Simulate 100 observations centered on r=0, and fit a linear model on either side.
```{r}
r <- runif(100, -3, 3)
treat <- as.numeric(r>0)
y <- fcntrl(r)*(1-treat) + ftreat(r)*treat
plot(r, y, col=1+(r>0))

rdfit <- glm( y ~ r*treat)
points(r, rdfit$fitted, pch=20, col=1+treat)
```

The CATE at r=0 is
```{r}
summary(rdfit)
```

The true value of 2 should be within 2 standard errors of the true value, 19 times out of 20 when you run this simulator code. 

<br> <br>
